---
title: "Assignment 10A: Sentiment Analysis"
author: "Emily El Mouaquite"
format: html
editor: visual
---

## Approach

1.  Reproduce the sentiment analysis found in chapter 2 of *Text Mining with R*.

2.  Extend the analysis using [this](https://www.kaggle.com/datasets/azminetoushikwasi/happiness-books-quotes-reviews-sentiments) dataset containing quotes from bestselling books about happiness.

3.  Utilize the Syuzhet lexicon for further analysis.

## Code Base

### Reproduction of Base Example

Example code from chapter 2 of *Text Mining with R*:

```{r}
library(tidytext)

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
```

```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

p_and_p_sentences$sentence[2]
```

```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

Works Cited

Silge, Julia, and David Robinson. “Chapter 2: Sentiment Analysis with Tidy Data.” *Text Mining with R: A Tidy Approach*, O’Reilly Media, Inc., 2017.

### Analysis Extension

This analysis can be extended through analyzing the Kaggle dataset [Bestseller Happiness Books 📚 Reviews, Quotes](https://www.kaggle.com/datasets/azminetoushikwasi/happiness-books-quotes-reviews-sentiments), which contains quotes from books about happiness.

```{r}
#read csv
happiness_books <- read.csv("data.csv")
head(happiness_books)
```
We will use the column favorite_quote in order to perform the sentiment analysis using the Syuzhet lexicon. Information on using this package that I used to inform the following sentiment analysis can be found [here](https://programminghistorian.org/en/lessons/sentiment-analysis-syuzhet).
```{r}
library(syuzhet)
#get sentiment scores 
syuzhet_scores <- get_sentiment(happiness_books$favorite_quote,
                                method = "syuzhet")
#add sentiment scores to the dataset 
happiness_books$sentiment <- syuzhet_scores
```

```{r}
ggplot(happiness_books, aes(x = sentiment)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of Sentiment Scores (Syuzhet)",
       x = "Sentiment Score",
       y = "Count")
```
The sentiment scores provided by the Syuzhet lexicon read as a lower score meaning a more negative sentiment, while a higher score alluding to a more positive sentiment. The above histogram is skewed to the left. This means that most of the sentiment scores for the quotes in the dataset are higher, or have a positive sentiment. This follows as the dataset contains quotes from books on happiness. There is one outlier with a sentiment score of almost -5.

## Conclusion
The results of the above sentiment analysis differ from the original example primarily because of the nature of the texts, as well as the sentiment method used. While the original analysis of novels by Jane Austen shows significant fluctuations in sentiment, the happiness quote dataset produces more consistently positive sentiment scores with less variation. Additionally, this analysis incorporates the Syuzhet sentiment lexicon, which calculates sentiment scores across entire text passages, resulting in smoother sentiment trends compared to the singular word level lexicon approaches used in the original example.